Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-24 arxiv/diff_cov.tex
wip

Interestingly, the prediction loss admits a bias-variance decomposition similar to the same covariates setting.
In both the bias and variance equations, the matrix inverse $\hat{\Sigma}^{-1}$ is a crucial quantity.
Intuitively, this matrix inverse combines the sample covariance matrix of both tasks, and it is not hard to see that the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by the sample sizes.
However, since the two tasks have different covariates, the scaling of $\hat{\Sigma}^{-1}$ not only depends on the tasks' sample sizes, but also depends on the covariate shift between the two tasks.
Building on the intuition, We introduce a key result that shows tight asymptotic convergence rate of the bias and variance as $p$ goes to infinity.
We define a key quantity $M \define \Sigma_1^{1/2}\Sigma_2^{-1/2}$ that captures the covariate shift between task $1$ and task $2$.
Let $U\Lambda V^\top$ denote the singular value decomposition of $M$, where $\Lambda\in\real^{p\times p}$ consists of the singular values of $M$, $\lambda_1, \lambda_2, \dots, \lambda_p$ in descending order.

\begin{theorem}
	Suppose that $X_1$ and $X_2$ are both random matrices that satisfy Assumption \ref{assume_rm}.
	With high probability over the randomness of $X_1$ and $X_2$, we have the following limits:
	\begin{enumerate}
		\item[i)]  The variance equation converges to the following asymptotic limit
			\begin{align}\label{lem_cov_shift_eq}
				\bigabs{\bigtr{\Sigma^{(2)} \bigbrace{\frac{\hat{\Sigma}^{-1}}{n_1 + n_2} - (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}}} \lesssim p^{-c_{\varphi}}\norm{\Sigma^{(2)}},
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following two equations:
			\begin{align}
				a_1 + a_2 = 1- \frac{1}{\rho_1 + \rho_2},\quad a_1 + \frac1{\rho_1 + \rho_2}\cdot \bigbrace{\frac{1}{p}\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{\rho_1}{\rho_1 + \rho_2}. \label{eq_a12extra}
			\end{align}
		\item[ii)] For any fixed vector $w\in\real^p$, the bias equation converges to the following asymptotic limit
			\be\label{def Pibias} \Pi_\bias \define  M \frac{a_3 M^{\top}M + (1 + a_4) \id_{p\times p}}{(a_1 M^{\top}M + a_2\id_{p\times p})^2} M^{\top}, \ee
				where $a_{3}$ and $a_4$ are the solutions of the following two equations:
			\begin{gather}\label{eq_a34extra}
				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
				= b_0 .
%		\left(\frac{n_1}{\hat a_1^2} -  \sum_{i=1}^p \frac{\hat \lambda_i^4   }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_4 -\left(\sum_{i=1}^p \frac{\hat \lambda_i^2  }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_3
%		= \sum_{i=1}^p \frac{\hat \lambda_i^2 }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }. \label{eq_a4}
			\end{gather}
				In the above equation, we denote $b_k$ as $\frac1{p}\sum_{i=1}^p {\lambda_i^{2k}} / { (\lambda_i^2 a_1 + a_2)^2  }$, for $k = 0, 1, 2$.
			Then, we have that
			\begin{align}\label{lem_cov_derv_eq}
				\bignorm{\Sigma_2^{1/2} \bigbrace{\hat{\Sigma}{n_1 + n_2}}^{-1} \Sigma_1^{1/2} w}^2
						= w^{\top} \Pi_\bias w + o(\|w\|^2),
			\end{align}
	\end{enumerate}
\end{theorem}

>>>>

Interestingly, the prediction loss admits a bias-variance decomposition similar to the same covariates setting.
In both the bias and variance equations, the matrix inverse $\hat{\Sigma}^{-1}$ is a crucial quantity.
Intuitively, this matrix inverse combines the sample covariance matrix of both tasks, and it is not hard to see that the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by the sample sizes.
However, since the two tasks have different covariates, the scaling of $\hat{\Sigma}^{-1}$ not only depends on the tasks' sample sizes, but also depends on the covariate shift between the two tasks.
Building on the intuition, We introduce a key result that shows tight asymptotic convergence rate of the bias and variance as $p$ goes to infinity.
We define a key quantity $M \define \Sigma_1^{1/2}\Sigma_2^{-1/2}$ that captures the covariate shift between task $1$ and task $2$.
Let $U\Lambda V^\top$ denote the singular value decomposition of $M$, where $\Lambda\in\real^{p\times p}$ consists of $M$'s singular values $\lambda_1, \lambda_2, \dots, \lambda_p$ as its diagonal entries in descending order.

\begin{theorem}\label{thm main RMT}
	Suppose that $X_1$ and $X_2$ are both random matrices that satisfy Assumption \ref{assume_rm}.
	With high probability over the randomness of $X_1$ and $X_2$, we have the following limits:
	\begin{enumerate}
		\item[i)]  The variance equation converges to the following asymptotic limit
			\begin{align}\label{lem_cov_shift_eq}
				\bigabs{\bigtr{\Sigma^{(2)} \bigbrace{(n_1 + n_2){\hat{\Sigma}^{-1}} - (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}}} \lesssim p^{-c_{\varphi}}\norm{\Sigma^{(2)}},
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following two equations:
			\begin{align}
				a_1 + a_2 = 1- \frac{1}{\rho_1 + \rho_2},\quad a_1 + \frac1{\rho_1 + \rho_2}\cdot \bigbrace{\frac{1}{p}\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{\rho_1}{\rho_1 + \rho_2}. \label{eq_a12extra}
			\end{align}
		\item[ii)] For any fixed vector $w\in\real^p$ with unit norm, the bias equation converges to the following asymptotic limit
			\begin{align}\label{def Pibias}
				\bigabs{w^{\top} \Sigma^{(1)} \bigbrace{(n_1+n_2)^2\hat{\Sigma}^{-1} \Sigma^{(2)} \hat{\Sigma}^{-1} - {\Sigma^{(2)}}^{-1/2} V {\frac{a_3 \Lambda^2 + (a_4 + 1)\id}{(a_1 \Lambda + a_2\id)^2}} V^{\top} {\Sigma^{(2)}}^{-1/2}} \Sigma^{(1)} w} \lesssim p^{-c_{\varphi}},
			\end{align}
				where $a_{3}$ and $a_4$ are the solutions of the following two equations with $b_k = \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}} {(\lambda_i^2 a_1 + a_2)^2}$, for $k = 0, 1, 2$:
			\begin{align}\label{eq_a34extra}
				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
				= b_0.
			\end{align}
	\end{enumerate}
\end{theorem}
